Will Kamala Harris will win the 2024 Presidential election

### 2024-11-04 paper/paper.qmd
discussion proffread

The paper is structured as follows: the Data section (@sec-data) covers our methodology and data collection process, detailing the variables used in the analysis. The Model section (@sec-model) introduces our Bayesian approach and includes related visualizations. The Results section (@sec-results) provides a model summary and predicted support percentages for each swing state. In the Discussion section (@sec-discussion), we interpret our findings, present final predictions, and discuss limitations, weaknesses and next steps.  We also give a final word about who will win the upcoming election. We also have additional appendices, where we talk about the methodology of the polls and how they compare. The second appendix talks about a survey we designed with reasons and disadvantage. The third appendix shows the Posterior Predictive check of our model.

@@

CONFIDENCE INTERVAL TABLE ??

@@

Here we will discuss how we believe each state wwill vote both base on data we collected, the models we created and the voting history of each of these swing states in past elections.

@@

Arizona is a very important swing state especially as the state has switches which party they voted for in in the last two elections with Trump winning 49% of the vote and Clinton with 45.5%, which considering Arizona is a swing state is a rather significant6 margin. While in the 2020 election Biden ended up winning Arizona with 49.4% of the vote and Trump 49.1%, which is a much smaller margin of victory and what makes Arizona so questionable in the 2024 election. We can see in @fig-supportAM that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls. Although, in the @fig-supportstate trump still has larger support in the state, and the largest polling difference in Trump's favour out of all the swing states polled @towin270.

We believe Arizona will vote: **Donald Trump**

@@

Considering the last two elections Georgia is very similar to Arizona, as in 2016 Trump won the state with 50.8% to Clinton's 45.6%. While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020, which was the first democratic victory for the state since 1992. Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, as the last victory for Biden was as a small margin, even a small percent of the population switching their vote could heavily impact the vote in 2024.

In @fig-supportGNN at the beginning of the polling data the support for Harris was relatively low, while it did start to grow the polling percentage somewhat plateaued at around 47.5% and does not show any signs of growing. Furthermore, in @fig-supportstate we can see that Trump has the edge over Harris considering all of these factors there is a very real possibility that Trump does win Georgia at the end of the day @towin270.

We believe Georgia will vote: **Donald Trump**

@@

We believe Nevada will vote: **Kamala Harris**

@@

We believe North Carolina will vote: **Donald Trump**

@@

We believe Michigan will vote for: **Kamala Harris**

@@

We believe Pennsylvania will vote: **Kamala Harris**

@@

We believe Wisconsin will vote: **Kamala Harris**

@@

As the election is quickly approachin there are not too many things we would be bale to consider for this election, nevertheless there are a few. First, While we did consider the historical voting paterns of each state, we have not done a rigorous analysis with that data, which we would be able to consider when analyzing a the topic in the future. As, poll results are inherently uncertain and are not absolute indicators of who will win the election, we could use confidence intervals in the future to demonstrate this uncertainty. Next, we could could add more variables into our analysis to create a stronger and more accurate model to help us predict the election with a higher degree of assurance. LAstly, we could consider intangibles more heavily in the future, including recent events, both local and worldwide, considering the socioeconomic status of voters and the voting patterns of certain ethnic groups and how the population and demographic has changed since the last election, as well as general sentiment towards the parties and their presidential candidates.

- integrate more context, and more vairables : Future studies could examine how state-level socioeconomic and political factors, such as unemployment rates, healthcare concerns, or local election issues, might influence voter preferences in each swing state. This would add depth to the model and align predictions more closely with real-world factors.

@@

# Pollster Methodology

@@

# Survey

@@

# Posterior Predictive Check {#sec-checks}

>>>>

The paper is structured as follows: the Data section (@sec-data) covers our methodology and data collection process, detailing the variables used in the analysis. The Model section (@sec-model) introduces our Bayesian approach and includes related visualizations. The Results section (@sec-results) provides a model summary and predicted support percentages for each swing state. In the Discussion section (@sec-discussion), we interpret our findings, present final predictions, and discuss limitations, weaknesses and next steps. We also give a final word about who will win the upcoming election. We also have additional appendices (@sec-app), where we talk about the methodology of the polls and how they compare. The second appendix talks about a survey we designed with reasons and disadvantage. The third appendix shows the Posterior Predictive check of our model.

@@

@@

In this section, we will analyze each state individually, considering our collected data, generated models, and the voting history of each state in previous elections to project likely outcomes.

@@

Arizona remains a pivotal swing state, particularly given its recent shifts in party alignment in the past two elections. In 2016, Trump won Arizona with 49% of the vote to Clinton's 45.5%, a relatively substantial margin for a swing state. However, in 2020, Biden narrowly won with 49.4% to Trump’s 49.1%, highlighting Arizona's competitiveness for the 2024 election, which is what makes Arizona such an important state.
As seen in @fig-supportAM, Harris's polling numbers were initially around 42%, but have significantly increased in the most recent polls. Despite this trend, @fig-supportstate indicates that Trump still has larger support in the state, with Arizona having the largest polling difference in his favor among the swing states (@towin270).

We project Arizona will vote for: **Donald Trump**

@@

Georgia's last two elections outcomes were very alike Arizona's outcomes. In 2016, Trump secured the state with 50.8% of the vote to Clinton's 45.6%. However, Biden won the 2020 election with a slim victory of 49.5% to Trump's 49.3%, which emphasizes Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020.This marked the state’s first Democratic victory since 1992. Georgia is a compelling state because most of the democratic vote comes from the Black community, although Harris has seen relatively lower support among Black men. Given Biden's last narrow win, even a minor shift in voter preferences could significantly impact the 2024 outcome.

As shown in @fig-supportGNN, Harris’s support rose slightly but plateaued around 47.5%  and does not show any signs of increasing. Furthermore, @fig-supportstate shows that Trump currently holds a slight advantage, suggesting a likely outcome in his favor (@towin270).

We project Georgia will vote for: **Donald Trump**

@@

We project Nevada will vote for: **Kamala Harris**

@@

We project North Carolina will vote for: **Donald Trump**

@@

We project Michigan will vote for: **Kamala Harris**

@@

We project Pennsylvania will vote for: **Kamala Harris**

@@

We project Wisconsin will vote for: **Kamala Harris**

@@

As the election is quickly approaching there are not too many things we would be bale to consider for this election, nevertheless there are a few. First, While we did consider the historical voting patterns of each state, we have not done a rigorous analysis with that data, which we would be able to consider when analyzing a the topic in the future. As, poll results are inherently uncertain and are not absolute indicators of who will win the election, we could use confidence intervals in the future to demonstrate this uncertainty. Next, we could could add more variables into our analysis to create a stronger and more accurate model to help us predict the election with a higher degree of assurance. Lastly, we could consider intangibles more heavily in the future, including recent events, both local and worldwide, considering the socioeconomic status of voters and the voting patterns of certain ethnic groups and how the population and demographic has changed since the last election, as well as general sentiment towards the parties and their presidential candidates.

@@

# Appendix {#sec-app}

## Pollster Methodology

@@

## Survey

@@

## Posterior Predictive Check {#sec-checks}
